---
title: "Assignment 1"
author: "Emma Colley"
date: "9/7/2020"
output: html_document

---

## Introduction

I am interested in studying rent burden in Massachusetts, so this data set aims to compare Gross Rent as a Percentage of Household Income to five other variables.

## Unit of Analysis

The unit of analysis is household-level data from the 2018 American Community Survey.

Initially, I wanted to study Gross Rent as a Percentage of Household Income (which is household-level data) in conjunction with person-level data to formulate household-level hypotheses. In order to do this I would need to derive a household-level variable from my person-level variables and collapse the data to the household level using the group_by and summarize functions. I aim to do this in Assignment 2 in order to bring in variables such as Race and Education Level. For now, I am sticking to household-level data only.

## Definition of Sample Population

The definition of the sample population is rental households in the state of Massachusetts who spend 30% or more of their household income on rent. I used 30% because this is the percentage HUD uses to define cost-burdened families.

## Definition of Variables

I selected four continuous variables and two categorical variables as shown below.

*Continuous*

1. Family Income (FINCP)
2. Gross Monthly Rent (GRNTP)
3. Gross Rent as Percentage of Household Income (GRPIP)
4. Number of People in Household (NP)
  
*Categorical*

1. Limited English Speaking Household (LNGI)
2. Grandparent Headed Household with no Parent (NPP)

## Loading Data

```{r load libraries, message = FALSE}
library(tidycensus)
library(tidyverse)
```

```{r load person-level microdata variabels}
person_vars_2018 <- pums_variables %>%
  distinct(year, survey, var_code, 
           var_label, data_type, level) %>%
  filter(level == "person", year == 2018, survey == "acs1")
```

```{r load household-level microdata variables}
hh_vars_2018 <- pums_variables %>%
  distinct(year, survey, var_code, 
           var_label, data_type, level) %>%
  filter(level == "housing", year == 2018, survey == "acs1")
```

## Defining Variables

In order to include only renters, I used the filter command to find rent payments greater than zero. I referenced my classmate Aron Lesser's code for this. I did the same for Family Income because I was initially receiving many incomes less than zero.

Finally, I used the filter command to show Gross Rent as a Percentage of Household Income greater than 29% and less than 101% to look at only rent burdened households.

There are 4,021 observations.


```{r load household-level data, results = 'hide'}
hh_data <- get_pums(variables = c("FINCP",
                                      "GRNTP",
                                      "GRPIP",
                                      "NP",
                                      "LNGI", 
                                      "NPP"),
                        state = "MA", 
                        year = 2018,
                        survey = "acs1",
                        recode = TRUE) %>%
  filter(GRNTP > 0) %>%
  filter(FINCP > 0) %>%
  filter(GRPIP > 29) %>%
  filter(GRPIP < 101) %>%
  select(FINCP, GRNTP, GRPIP, NP, LNGI_label, NPP_label)
```

## First Ten Observations

There are 4,021 observations, here are 10.

```{r}
hh_data
```


## Writing CSV File

```{r save dataframes as csv}
write_csv(hh_data, "renters_massachusetts.csv")
```

