---
title: "Week 4 Assignment"
author: "Emma Colley"
date: "10/01/2020"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: lumen
    code_floating: hide
---

# Topic

Rent Burden in Massachusetts

*Examining Gross Rent as a Percentage of Household Income against four other variables*

# Variables

*Continuous*

1. Age of Individual (AGEP)
2. Number of People in Household (NP)
3. Annual Household Income (HINCP)
  
*Categorical*

1. Race/Ethnicity (RAC1P)
2. Highest Level of Education Received (SCHL)

## Loading Libraries

```{r load libraries, message = FALSE}
library(tidycensus)
library(tidyverse)
library(ggplot2)
library(ggthemes)
library(dbplyr)
library(dplyr)
```

## Loading Data

I am going to re-code some data into fewer categories. Please note that for the purposes of this exercise I have grouped American Indian, Alaska Native, Native Hawaiian, Pacific Islander and Native tribes into one category. I understand I would want to be more critical about grouping racial identities in future analysis. I have also grouped education levels together, including Bachelors, Masters and Doctorate degree recipients. I understand there are major differences between the three and would break this down under other circumstances. 

```{r, message = FALSE}
person_data <- read_csv("rent_burden_massachusetts_v3.csv") %>%

mutate(EDU = case_when(
    SCHL_label=="1 or more years of college credit, no degree"~"HS Diploma",
    SCHL_label=="12th grade - no diploma"~"HS Diploma",
    SCHL_label=="Associate's degree"~"Assoc. Degree",
    SCHL_label=="Bachelor's degree"~"Bach Degree or Higher",
    SCHL_label=="Doctorate degree"~"Bach Degree or Higher",
    SCHL_label=="GED or alternative credential"~"HS Diploma",
    SCHL_label=="Grade 10"~"1No HS",
    SCHL_label=="Grade 11"~"1No HS",
    SCHL_label=="Grade 1"~"1No HS",
    SCHL_label=="Grade 2"~"1No HS",
    SCHL_label=="Grade 3"~"1No HS",
    SCHL_label=="Grade 4"~"1No HS",
    SCHL_label=="Grade 5"~"1No HS",
    SCHL_label=="Grade 6"~"1No HS",
    SCHL_label=="Grade 7"~"1No HS",
    SCHL_label=="Grade 8"~"1No HS",
    SCHL_label=="Grade 9"~"1No HS",
    SCHL_label=="Kindergarten"~"1No HS",
    SCHL_label=="Master's degree"~"Bach Degree or Higher",
    SCHL_label=="No schooling completed"~"1No HS",
    SCHL_label=="Nursery school, preschool"~"1No HS",
    SCHL_label=="Professional degree beyond a bachelor's degree"~"Bach Degree or Higher",
    SCHL_label=="Regular high school diploma"~"HS Diploma",
    SCHL_label=="Some college, but less than 1 year"~"HS Diploma"),

RACE = case_when(
    RAC1P_label=="American Indian alone"~"Native Am.",
    RAC1P_label=="American Indian and Alaska Native tribes specified; or American Indian or Alaska Native, not specified and no other races"~"Native Am.",
    RAC1P_label=="Asian alone"~"Asian",
    RAC1P_label=="Black or African American alone"~"Black",
    RAC1P_label=="Native Hawaiian and Other Pacific Islander alone"~"Native Am.",
    RAC1P_label=="Some Other Race alone"~"Other Race", 
    RAC1P_label=="Two or More Races"~"Multiracial",   
    RAC1P_label=="White Alone"~"1White"))
```


```{r}
model <- lm(GRPIP ~ EDU + RACE + AGEP + NP, 
            data = person_data)

summary(model)
```
## Intrepretation

The multiple R-squared value for this model is 0.03515, which tells us that this model predicts about 4% of the variation in gross rent as a percentage of household income in this dataset.

*Education Level Analysis*

I am comparing education levels to the least amount of education: No High School Diploma.

A household with someone with a high school diploma pays, on average, 2.2 percentage points less on rent as a percentage of household income than a household with no high school degree, controlling for the other variables in the model. This is statistically significant at a 95-percent confidence level.

A household with someone with an associate's degree pays, on average, 2.6 percentage points less on rent as a percentage of household income than a household with no high school degree, controlling for the other variables in the model. This is not statistically significant at a 95-percent confidence level.

A household with someone with a Bachelor's degree or higher pays, on average, 7.5 percentage points less on rent as a percentage of household income than a household with no high school degree, controlling for the other variables in the model. This is statistically significant at a 95-percent confidence level.

*Race Analysis*

I am comparing races to the most common race in my sample set: White alone.

A household with at least one person who is black or African American pays, on average, 0.3 percentage points less on rent as a percentage of household income than an all white household, controlling for the other variables in the model. This is not statistically significant at a 95-percent confidence level.

A household with at least one person who is biracial or multiracial pays, on average, 0.7 percentage points less on rent as a percentage of household income than an all white household, controlling for the other variables in the model. This is not statistically significant at a 95-percent confidence level.

A household with at least one person who is Native American, Pacific Islander, Native Alaskan or Native Hawaiian pays, on average, 8.0 percentage points less on rent as a percentage of household income than an all white household, controlling for the other variables in the model. This is statistically significant at a 95-percent confidence level.

A household with at least one person who is some other race pays, on average, 0.3 percentage points less on rent as a percentage of household income than an all white household, controlling for the other variables in the model. This is not statistically significant at a 95-percent confidence level.

*Other Analysis*

An increase in age by one year is associated with an average decrease in rent as a percentage of household income by 0.03 percentage points, on average, controlling for the other variables in the model. This is not statistically significant at a 95-percent confidence level.

As you increase the household size by one person, the rent as a percentage of household income decreases by 1.8 percentage points, on average, controlling for the other variables in the model. This is statistically significant at a 95-percent confidence level
