---
title: "Assignment 1"
author: "Emma Colley"
date: "9/7/2020"
output: html_document

---

# Rent Burden in Massachusetts

This data set aims to compare Gross Rent as a Percentage of Household Income to other variables. I have more than five variables and hope to refine them in the coming weeks.

## Unit of Analysis

The unit of analysis is person-level data from the 2018 American Community Survey.

## Definition of Sample Population

The definition of the sample population is residents of the state of Massachusetts age 18 or older.

## Definition of Variables

*Continuous*

1. Family Income (FINCP)
2. Gross Monthly Rent (GRNTP)
3. Gross Rent as Percentage of Household Income (GRPIP)
4. Age (AGEP)
  
*Categorical*

1. Sex (SEX)
2. Education Level (SCHL)
3. Race (RAC1P)

## Loading Data

```{r load libraries}
library(tidycensus)
library(tidyverse)
```

```{r load person-level microdata variabels}
person_vars_2018 <- pums_variables %>%
  distinct(year, survey, var_code, 
           var_label, data_type, level) %>%
  filter(level == "person", year == 2018, survey == "acs1")
```

```{r load household-level microdata variables}
hh_vars_2018 <- pums_variables %>%
  distinct(year, survey, var_code, 
           var_label, data_type, level) %>%
  filter(level == "housing", year == 2018, survey == "acs1")
```

## Defining Variables

```{r load person-level data, results = 'hide'}
person_data <- get_pums(variables = c("FINCP",
                                      "GRNTP",
                                      "GRPIP",
                                      "SEX",
                                      "AGEP", 
                                      "SCHL", 
                                      "RAC1P"),
                        state = "MA", 
                        year = 2018, 
                        survey = "acs1",
                        recode = TRUE) %>%
  filter(GRNTP > 0) %>%
  filter(FINCP > 0) %>%
  filter(AGEP > 17) %>%
  select(FINCP, GRNTP, GRPIP, SEX_label, AGEP, SCHL_label, RAC1P_label)
```

## First Ten Observations
```{r}
person_data
```


## Writing CSV File

```{r save dataframes as csv}
write_csv(person_data, "renters_massachusetts.csv")
```

