---
title: "Week 3 Assignment"
author: "Emma Colley"
date: "9/21/2020"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: lumen
    code_floating: hide
---

# Redefining Variables

While I've kept my overall topic of rent burden in Massachusetts, I've made some small updates since last week. Instead of including people who spend 30% - 100% of their gross household income on monthly rent, I am now including all people who spend 0-100% of their gross household income on rent. I think this will give more depth to my findings below. Also, when examining the number of people in a household, I am capping the number at 10 people. There were a few outliers with 11, 12 and 13 people.

*Continuous*

1. Age of Individual (AGEP)
2. Number of People in Household (NP)
3. Annual Household Income (HINCP)
4. Gross Rent as Percentage of Household Income (GRPIP)
  
*Categorical*

1. Race/Ethnicity (RAC1P)
2. Highest Level of Education Received (SCHL)

## Loading Libraries

```{r load libraries, message = FALSE}
library(tidycensus)
library(tidyverse)
library(ggplot2)
library(ggthemes)
```

## Loading Data

I am going to re-code the education level data to break it into fewer categories. I'll attempt to use this later.

```{r, message = FALSE}
person_data <- read_csv("rent_burden_massachusetts_v3.csv") %>%
mutate(EDU = case_when(
    SCHL_label=="1 or more years of college credit, no degree"~"GED or High School Diploma",
    SCHL_label=="12th grade - no diploma"~"No High School Diploma",
    SCHL_label=="Associate's degree"~"Associate's Degree",
    SCHL_label=="Bachelor's degree"~"Bachelor's Degree or Higher",
    SCHL_label=="Doctorate degree"~"Bachelor's Degree or Higher",
    SCHL_label=="GED or alternative credential"~"GED or High School Diploma",
    SCHL_label=="Grade 10"~"No High School Diploma",
    SCHL_label=="Grade 11"~"No High School Diploma",
    SCHL_label=="Grade 1"~"No High School Diploma",
    SCHL_label=="Grade 2"~"No High School Diploma",
    SCHL_label=="Grade 3"~"No High School Diploma",
    SCHL_label=="Grade 4"~"No High School Diploma",
    SCHL_label=="Grade 5"~"No High School Diploma",
    SCHL_label=="Grade 6"~"No High School Diploma",
    SCHL_label=="Grade 7"~"No High School Diploma",
    SCHL_label=="Grade 8"~"No High School Diploma",
    SCHL_label=="Grade 9"~"No High School Diploma",
    SCHL_label=="Kindergarten"~"No High School Diploma",
    SCHL_label=="Master's degree"~"Bachelor's Degree or Higher",
    SCHL_label=="No schooling completed"~"No High School Diploma",
    SCHL_label=="Nursery school, preschool"~"No High School Diploma",
    SCHL_label=="Professional degree beyond a bachelor's degree"~"Bachelor's Degree or Higher",
    SCHL_label=="Regular high school diploma"~"GED or High School Diploma",
    SCHL_label=="Some college, but less than 1 year"~"GED or High School Diploma"))
```

# Relationships between continuous variables that I'm interested In

Here are Pearson's tests and accompanying figures of Gross Rent as a Percentage of Household Income compared to other variables.


## Age of Induvidual and Gross Rent as a Percentage of their Household Income

This correlation is weak but statistically significant. There is a slightly positive correlation at a 95% confidence level. 

```{r, message=FALSE, results='hide'}
correlation <- cor.test(person_data$AGEP, person_data$GRPIP)

correlation
```
```{r, message=FALSE, results='hide'}
ggplot(person_data, aes(x = AGEP, y = GRPIP)) +
  geom_point(size = 0.6, alpha = 0.25, position = "jitter") +
  geom_smooth() +
  theme_bw() +
  scale_x_continuous(name = "Age of Induvidual", breaks = seq(18, 100, by = 10)) +
  scale_y_continuous(name = "Gross monthly rent as a Percentage of Household Income",
                     breaks = seq(0, 100, by = 10),
                     labels = paste(prettyNum(seq(0, 100, by = 10)), "%")) +
  annotate(geom = "text", x = 18, y = 105, 
           label = paste("Correlation (95-percent confidence):",
                         prettyNum(correlation$conf.int[1], digits = 2), 
                         "to",
                         prettyNum(correlation$conf.int[2], digits = 2)),
           hjust = 0)
```


## Number of People in Household and Gross Rent as a Percentage of their Household Income

This correlation is weak but statistically significant. There is a negative correlation at a 95% confidence level. 

```{r, message=FALSE, results='hide'}
correlation <- cor.test(person_data$NP, person_data$GRPIP)

correlation
```
```{r, message=FALSE, results='hide'}
ggplot(person_data, aes(x = NP, y = GRPIP)) +
  geom_point(size = 0.6, alpha = 0.25, position = "jitter") +
  geom_smooth() +
  theme_bw() +
  scale_x_continuous(name = "Number of People in Household", breaks = seq(1, 10, by = 1)) +
  scale_y_continuous(name = "Gross monthly rent as a Percentage of Household Income",
                     breaks = seq(0, 100, by = 10),
                     labels = paste(prettyNum(seq(0, 100, by = 10)), "%")) +
  annotate(geom = "text", x = 1, y = 105, 
           label = paste("Correlation (95-percent confidence):",
                         prettyNum(correlation$conf.int[1], digits = 2), 
                         "to",
                         prettyNum(correlation$conf.int[2], digits = 2)),
           hjust = 0)
```


## Household Income and Gross Rent as a Percentage of Household Income

This correlation is only moderate (around 0.5) but statistically significant. There is a negative correlation at a 95% confidence level. 

```{r, message=FALSE, results='hide'}
correlation <- cor.test(person_data$HINCP, person_data$GRPIP)

correlation
```

```{r, message=FALSE, results='hide'}
ggplot(person_data, aes(x = HINCP, y = GRPIP)) +
  geom_point(size = 0.6, alpha = 0.25) +
  geom_smooth() +
  theme_bw() +
  scale_x_continuous(name = "Household Income") +
  scale_y_continuous(name = "Gross monthly rent as a Percentage of Household Income",
                     breaks = seq(0, 100, by = 10),
                     labels = paste(prettyNum(seq(0, 100, by = 10)), "%")) +
  annotate(geom = "text", x = 0, y = 105, 
           label = paste("Correlation (95-percent confidence):",
                         prettyNum(correlation$conf.int[1], digits = 2), 
                         "to",
                         prettyNum(correlation$conf.int[2], digits = 2)),
           hjust = 0)
```

# Other relationships between two continuous variables

Here are Pearson's tests of other continuous variables that aren't GRPIP.


## Age of Induvidal and Number of People in their Household

This correlation is weak (around 0.3) and statistically significant. There is a negative correlation at a 95% confidence level. 

```{r, message=FALSE, results='hide'}
correlation <- cor.test(person_data$AGEP, person_data$NP)

correlation
```


## Age of Induvidal and their Household Income

This correlation is weak but statistically significant. There is a negative correlation at a 95% confidence level. 

```{r, message=FALSE, results='hide'}
correlation <- cor.test(person_data$AGEP, person_data$HINCP)

correlation
```


# Relationships between a continuous variable and a binary variable that I'm interested in

Here are Two Sample t-tests and accompanying figures for Gross Rent as a Percentage of Household Income and binary variables.


## Gross Rent as Percentage of Household Income and Education Level

Are people with a Bachelor's Degree less likely to live in rent burdened households?

There is a significant relationship between these two variables. We can be 95% confident that people in this sample set with a four-year college degree (Bachelor's or higher) live in households that pay 5% - 6% less of their income on rent than those without a Bachelor's degree. It looks like many people with a college degree live in households that pay less than 20% of their income on rent.

```{r, message=FALSE, results='hide'}
difference = t.test(GRPIP ~ EDU == "Bachelor's Degree or Higher",
                    data = person_data)

difference
```
```{r, message=FALSE, results='hide'}
ggplot(person_data, aes(x = EDU == "Bachelor's Degree or Higher", y = GRPIP)) +
  geom_violin() +
  theme_bw() +
  scale_x_discrete(name = "Does Someone in the Household Have a Bachelor's Degree or Higher?",
                   labels = c("No", "Yes")) +
   scale_y_continuous(name = "Gross monthly rent as Percentage of Household Income", 
                     breaks = seq(0, 100, by = 10),
                     labels = paste(prettyNum(seq(0, 100, by = 10)), "%")) +
  annotate(geom = "text", x = 1.5, y = 90, size = 3,
           label = paste("Difference in means\n(95% confidence):\n",
                         prettyNum(abs(difference$conf.int[2]), 
                                   digits = 0), "% to ",
                         prettyNum(abs(difference$conf.int[1]), 
                                   digits = 0), "%",
                         sep = "")) +
  annotate(geom = "text", x = 1, y = difference$estimate[1], size = 3,
           label = paste("Mean (95% confidence):\n",
                         prettyNum(difference$estimate[1], digits = 0), "%")) +
  annotate(geom = "text", x = 2, y = difference$estimate[2], size = 3,
           label = paste("Mean (95% confidence):\n",
                         prettyNum(difference$estimate[2], digits = 0), "%"))
```

## Gross Rent as Percentage of Household Income and Race

There is a significant relationship between these two variables. We can be 95% confident that households with at least one person who is white that pay 2% - 3% less of their income on rent than non-white households. Again, many more household with at least one white person pay less than 20% of their income on rent. I thought this would be slightly more but I am not defining entirely white vs. entirely non-white households at this time.

```{r, message=FALSE, results='hide'}
difference = t.test(GRPIP ~ RAC1P_label == "White alone",
                    data = person_data)

difference
```
```{r, message=FALSE, results='hide'}
ggplot(person_data, aes(x = RAC1P_label == "White alone", y = GRPIP)) +
  geom_violin() +
  theme_bw() +
  scale_x_discrete(name = "Is someone in the household white?",
                   labels = c("No", "Yes")) +
   scale_y_continuous(name = "Gross monthly rent as Percentage of Household Income", 
                     breaks = seq(0, 100, by = 10),
                     labels = paste(prettyNum(seq(0, 100, by = 10)), "%")) +
  annotate(geom = "text", x = 1.5, y = 90, size = 3,
           label = paste("Difference in means\n(95% confidence):\n",
                         prettyNum(abs(difference$conf.int[2]), 
                                   digits = 0), "% to ",
                         prettyNum(abs(difference$conf.int[1]), 
                                   digits = 0), "%",
                         sep = "")) +
  annotate(geom = "text", x = 1, y = difference$estimate[1], size = 3,
           label = paste("Mean (95% confidence):\n",
                         prettyNum(difference$estimate[1], digits = 0), "%")) +
  annotate(geom = "text", x = 2, y = difference$estimate[2], size = 3,
           label = paste("Mean (95% confidence):\n",
                         prettyNum(difference$estimate[2], digits = 0), "%"))
```



# Relationship between a continuous variable and a categorical variable (three or more levels) that I'm interested in

here are ANOVA test and accompanying figures for Gross Rent as a Percentage of Household Income and categorical variables.


## Gross Rent as Percentage of Household Income and Race

There is a significant relationships between an individual's race and their rent as a gross percentage of their household income. 

```{r, message=FALSE, results='hide'}
anova <- aov(GRPIP ~ RAC1P_label, data = person_data)

summary(anova)
```

```{r, message=FALSE, results='hide'}
differences <- TukeyHSD(anova)

as_tibble(cbind(pair = row.names(differences$RAC1P),
                differences$RAC1P))
```

```{r, message=FALSE, results='hide'}
ggplot(person_data, aes(x = RAC1P_label, y = GRPIP)) +
  geom_boxplot() +
  theme_bw() +
  scale_x_discrete(name = "Race", labels = c("American\nIndian", 
                            "American\nIndian\nAlaska Native\ntribes", 
                            "Asian",
                            "Black or\nAfrican\nAmerican",
                            "Native\nHawaiian\nPacific\nIslander",
                            "Some\nOther Race",
                            "Two or\nMore Races",
                            "White")) +

  scale_y_continuous(name = "Gross Monthly Rent as Percentage of Household Income", 
                     breaks = seq(0, 100, by = 10),
                     labels = paste(prettyNum(seq(0, 100, by = 10)), "%")) 
```

# Other relationships between continuous variables and categorical variables (three or more levels)

Here are ANOVA tests for continuous and categorical variables that don't include GRPIP.


## Race of an Induvidual and Their Household Income

There is a significant relationships between an individual's race and their household income. 

```{r}
anova <- aov(HINCP ~ RAC1P_label, data = person_data)

summary(anova)
```
```{r}
differences <- TukeyHSD(anova)

as_tibble(cbind(pair = row.names(differences$RAC1P),
                differences$RAC1P))
```


## Level of Education of an Induvidual and Their Household Income

There is a significant relationships between an individual's level of education and their household income. 

```{r}
anova <- aov(HINCP ~ EDU, data = person_data)

summary(anova)
```
```{r}
differences <- TukeyHSD(anova)

as_tibble(cbind(pair = row.names(differences$EDU),
                differences$EDU))
```


# Relationship between two categorical variables

Here is a Chi-Square test.

## Race of An Induvidual and their Level of Education

For this data set of residents of Massachusetts, is there a relationship between an individuals race and their education level?

Note: I am receiving a notification that "Chi=squared approximation may be incorrect." My classmate Megan Willis-Jackson explained to me this is because there are some race categories with very few observations. 

There is a statistically significant relationship between an individual's race and their level of education.

```{r}
chi_sq <- chisq.test(person_data$RAC1P_label, person_data$SCHL_label == "Master's degree") 

chi_sq
```
```{r}
chi_sq$observed
```
```{r}
chi_sq$expected
```

